Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-10-19 prop_letter.tex
prop_letter.tex \section{Gleason} Updating...

\noindent When generalizing to finite precision QIVPMs, we also want
to seek a simple rule to construct them, and to understand which states
are consistent with them. However, if the measurement intervals are
too imprecise and the QIVPM are defined so wildly, the results we
get might contradict to each other. This contradicting results would
exclude any possible state from being consistent with them.

@@

\begin{definition}[THOS rule]\label{def:THOS}Given a collection
of intervals $\mathscr{I}$ including $\imposs$ and $\necess$, a
function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$ is called
a THOS rule if $\iota:\left[0,1\right]\rightarrow\mathscr{I}$ is
subject to the following constraints:\begin{subequations}\label{eq:iota-constraints}
\begin{eqnarray}
 &  & \iota(0)=\imposs\,,\\
 &  & \iota(1)=\necess\,,\\
 &  & \iota\left(1-x\right)=\necess-\iota\left(x\right)\,,
\end{eqnarray}
\end{subequations}and satisfies that for any three numbers~$x_{0}$,
$x_{1}$, and $x_{2}\in\left[0,1\right]$, their sum~$x_{0}+x_{1}+x_{2}\in\left[0,1\right]$
implies
\begin{equation}
\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\subseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)\,.\label{eq:iota-Inclusion}
\end{equation}
\end{definition}

@@

\begin{thm}[The equivalence among THOS rules and IVPMs]\label{thm:iota-statements}The
following three statements are equivalent:
\begin{enumerate}
\item \label{enu:iota-subject-to}The function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
is a THOS rule.
\item \label{enu:iota-mu-CIVPM}The composited function~$\iota\circ\mu:\events_{C}\rightarrow\mathscr{I}$
is always a classical IVPM for any classical probability measure~$\mu:\events_{C}\rightarrow\left[0,1\right]$.
\item \label{enu:iota-mu-QIVPM}The composited function~$\iota\circ\mu:\events\rightarrow\mathscr{I}$
is always a QIVPM for any quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
on any Hilbert space~$\Hilb$.
\end{enumerate}
\end{thm}

As we have discussed, Statement~\ref{enu:iota-subject-to} implies
\ref{enu:iota-mu-QIVPM}. The rest of the proof is to prove that if
$\iota$ is not a THOS rule, then $\iota\circ\mu$ might not be a
QIVPM. If $\iota$ doesn't satisfy Eq.~(\ref{eq:iota-Inclusion}),
we can find three particular numbers~$x_{0}$, $x_{1}$, and $x_{2}\in\left[0,1\right]$
such that $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$, but $\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\nsubseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)$.
Consider the state $\rho=x_{0}\proj{0}+x_{1}\proj{1}+x_{2}\proj{2}+\left(1-x_{0}-x_{1}-x_{2}\right)\proj{3}$.
Then, the function~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}:\events\rightarrow\mathscr{I}$
fails to satisfy Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
\begin{eqnarray*}
 &  & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{1}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{2}\right)\right)\\
 & \nsubseteq & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{1}+\proj{2}\right)\right)\,,
\end{eqnarray*}
and cannot be a QIVPM. Therefore, $\iota$ is a THOS rule if and only
if $\iota\circ\mu$ is always a QIVPM. Similarly, Statement~\ref{enu:iota-subject-to}
and \ref{enu:iota-mu-CIVPM} are equivalent as well~\citep{HOSTunpublished}.

@@

Given a state~$\rho'$ consistent with $\iota\circ\mu_{\rho}^{\mathrm{B}}$,
we have $\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)\in\iota\left(\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)\right)$
for any 1-dimensional projector~$\proj{\psi}$. Since the maximum
length of the intervals in $\mathscr{I}$ is $\delta$, it is also
the upper bound of the difference $\left|\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)-\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)\right|=\left|\melem{\psi}{\rho-\rho'}{\psi}\right|$.
Since $\rho-\rho'$ is Hermitian, $\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{\rho-\rho'}{\psi}\right|$
is the maximum absolute value of the eigenvalues of $\rho-\rho'$~\citep{544199},
and equal to $\left\Vert \rho-\rho'\right\Vert $~\citep{GolubVanLoan1996,Foucart2012}.
Hence, $\left\Vert \rho-\rho'\right\Vert \le\delta$.

@@

Since $\iota$ maps to intervals, we can decompose it into two functions:
its left-end and right-end~$\left[\iotal\left(x\right),\iotar\left(x\right)\right]=\iota\left(x\right)$.
By Eq.~(\ref{eq:iota-Inclusion}), the left-end function~$\iotal:\left[0,1\right]\rightarrow\left[0,1\right]$
is Wright-convex, i.e., $\iotal\left(x_{0}+x_{1}+x_{2}\right)+\iotal\left(x_{2}\right)\ge\iotal\left(x_{0}+x_{2}\right)+\iotal\left(x_{1}+x_{2}\right)$~\citep{Wright1954,RobertsVarberg1973,PecaricTong1992}.
It is well-known that a Wright-convex function is midpoint-convex
or Jensen-convex~\citep{Klee_1956,Kenyon_1956,RobertsVarberg1973,PecaricTong1992},
i.e., we have
\begin{equation}
\frac{\iotal\left(x_{0}\right)+\iotal\left(x_{1}\right)}{2}\ge\iotal\left(\frac{x_{0}+x_{1}}{2}\right)\,.
\end{equation}
Together with the fact that $\iotal$ maps to a bounded interval~$\left[0,1\right]$,
the left-end function~$\iotal$ must be continuous on the unit open
interval~$\left(0,1\right)$~\citep{Jensen_1906,RobertsVarberg1973}.
Therefore, either $\iota$ maps every number to the same interval,
or the number of intervals that $\iota$ maps to must be uncountable.

>>>>

\noindent For general QIVPMs mapping to finitely precise intervals,
we also want to seek a simple rule to construct them, and to understand
which states are consistent with them. However, if measurement intervals
are too imprecise and QIVPMs are defined so wildly, the results we
get might contradict to each other. These contradicting results would
exclude any possible state from being consistent with the given QIVPM.

@@

\begin{definition}[THOS rule]\label{def:THOS}Given a collection
of intervals $\mathscr{I}$ including $\imposs$ and $\necess$, a
function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$ is called
a THOS rule if $\iota$ is subject to the following constraints:\begin{subequations}\label{eq:iota-constraints}
\begin{eqnarray}
 &  & \iota(0)=\imposs\,,\\
 &  & \iota(1)=\necess\,,\\
 &  & \iota\left(1-x\right)=\necess-\iota\left(x\right)\,,
\end{eqnarray}
\end{subequations}and satisfies that for any three numbers~$x_{0}$,
$x_{1}$, and $x_{2}\in\left[0,1\right]$, their sum~$x_{0}+x_{1}+x_{2}\in\left[0,1\right]$
implies
\begin{equation}
\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\subseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)\,.\label{eq:iota-Inclusion}
\end{equation}
\end{definition}

@@

\begin{thm}[The equivalence among THOS rules and IVPMs]\label{thm:iota-statements}The
following three statements are equivalent:
\begin{enumerate}
\item \label{enu:iota-subject-to}A function~$\iota:\left[0,1\right]\rightarrow\mathscr{I}$
is a THOS rule.
\item \label{enu:iota-mu-CIVPM}The composited function~$\iota\circ\mu:\events_{C}\rightarrow\mathscr{I}$
is always a classical IVPM for any classical probability measure~$\mu:\events_{C}\rightarrow\left[0,1\right]$.
\item \label{enu:iota-mu-QIVPM}The composited function~$\iota\circ\mu:\events\rightarrow\mathscr{I}$
is always a QIVPM for any quantum probability measure~$\mu:\events\rightarrow\left[0,1\right]$
on any Hilbert space~$\Hilb$.
\end{enumerate}
\end{thm}

As we have discussed, Statement~\ref{enu:iota-subject-to} implies
\ref{enu:iota-mu-QIVPM}. The rest of the proof is to prove that if
$\iota$ is not a THOS rule, then $\iota\circ\mu$ might not be a
QIVPM. If $\iota$ doesn't satisfy Eq.~(\ref{eq:iota-Inclusion}),
we can find three particular numbers~$x_{0}$, $x_{1}$, and $x_{2}\in\left[0,1\right]$
such that $x_{0}+x_{1}+x_{2}\in\left[0,1\right]$, but $\iota\left(x_{0}+x_{1}+x_{2}\right)+\iota\left(x_{2}\right)\nsubseteq\iota\left(x_{0}+x_{2}\right)+\iota\left(x_{1}+x_{2}\right)$.
Consider the state $\rho=x_{0}\proj{0}+x_{1}\proj{1}+x_{2}\proj{2}+\left(1-x_{0}-x_{1}-x_{2}\right)\proj{3}$.
Then, the function~$\iota\circ\mu_{\rho}^{\mathrm{\mathrm{B}}}$
fails to satisfy Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
\begin{eqnarray*}
 &  & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{1}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{2}\right)\right)\\
 & \nsubseteq & \iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{0}+\proj{2}\right)\right)+\iota\left(\mu_{\rho}^{\mathrm{\mathrm{B}}}\left(\proj{1}+\proj{2}\right)\right)\,,
\end{eqnarray*}
and cannot be a QIVPM. Therefore, $\iota$ is a THOS rule if and only
if $\iota\circ\mu$ is always a QIVPM. Similarly, Statement~\ref{enu:iota-subject-to}
and \ref{enu:iota-mu-CIVPM} are equivalent as well~\citep{HOSTunpublished}.

@@

Given a state~$\rho'$ consistent with $\iota\circ\mu_{\rho}^{\mathrm{B}}$,
we have $\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)\in\iota\left(\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)\right)$
for any 1-dimensional projector~$\proj{\psi}$. Since the maximum
length of the intervals in $\mathscr{I}$ is $\delta$, it is also
the upper bound of the difference 
\[
\left|\mu_{\rho'}^{\mathrm{B}}\left(\proj{\psi}\right)-\mu_{\rho}^{\mathrm{B}}\left(\proj{\psi}\right)\right|=\left|\melem{\psi}{\rho-\rho'}{\psi}\right|\le\delta\,.
\]
Since $\rho-\rho'$ is Hermitian, $\max_{\ip{\psi}{\psi}=1}\left|\melem{\psi}{\rho-\rho'}{\psi}\right|$
is the maximum absolute value of the eigenvalues of $\rho-\rho'$~\citep{544199},
and equal to $\left\Vert \rho-\rho'\right\Vert $~\citep{GolubVanLoan1996,Foucart2012}.
Hence, $\left\Vert \rho-\rho'\right\Vert \le\delta$.

@@

Since $\iota$ maps to intervals, we can decompose it into two functions:
its left-end and right-end $\left[\iotal\left(x\right),\iotar\left(x\right)\right]=\iota\left(x\right)$.
By Eq.~(\ref{eq:iota-Inclusion}), the left-end function $\iotal:\left[0,1\right]\rightarrow\left[0,1\right]$
is Wright-convex, i.e., $\iotal\left(x_{0}+x_{1}+x_{2}\right)+\iotal\left(x_{2}\right)\ge\iotal\left(x_{0}+x_{2}\right)+\iotal\left(x_{1}+x_{2}\right)$~\citep{Wright1954,RobertsVarberg1973,PecaricTong1992}.
It is well-known that a Wright-convex function is midpoint-convex
or Jensen-convex~\citep{Klee_1956,Kenyon_1956,RobertsVarberg1973,PecaricTong1992},
i.e., we have
\begin{equation}
\frac{\iotal\left(x_{0}\right)+\iotal\left(x_{1}\right)}{2}\ge\iotal\left(\frac{x_{0}+x_{1}}{2}\right)\,.
\end{equation}
Together with the fact that $\iotal$ maps to a bounded interval~$\left[0,1\right]$,
the left-end function~$\iotal$ must be continuous on the unit open
interval~$\left(0,1\right)$~\citep{Jensen_1906,RobertsVarberg1973}.
Therefore, either $\iota$ maps every number to the same interval,
or the number of intervals that $\iota$ maps to must be uncountable.
